---
title: "p8105 homework1"
author: "Shuwei Liu sl4471"
date: 2018-09-19
output: github_document
---

```{r setup, include = FALSE}
library(tidyverse)
```

This is the first homework for p8105. 

# Problem 1

## 1.1 Create a data frame

First, create a data frame containing four variables as the problem 1 asked.

```{r problem1_data_frame}
set.seed(1)

hw1_df = tibble(
  vec_numeric = runif(10, min = 0, max = 5),
  vec_logical = vec_numeric > 2,
  vec_char = c("hello", "data", "science", "my", "friend", "I", "can", "figure", "it", "out"),
  vec_factor = factor(c("apple", "banana", "orange", "grape", "peach", "lemon", "stawberry", "blueberry", "pear", "watermelon"))
)

hw1_df
```

## 1.2 Take the mean of variables

Next, I'll try to take the mean of each variables.

The mean of variables "vec_numeric" is `r mean(hw1_df$vec_numeric)`.
The mean of variables "vec_logical" is `r mean(hw1_df$vec_logical)`.
The mean of variables "vec_char" is `r mean(hw1_df$vec_char)`.
The mean of variables "vec_factor" is `r mean(hw1_df$vec_factor)`.

As above, obviously, only the variable "vec_numeric" and "vec_logical" have mean since the other variables are words that cannot be calculated.

## 1.3 Apply the "as." function to the other variables

```{r apply as.numeric function, results='hide', warning=FALSE}
as.numeric(hw1_df$vec_logical)
as.numeric(hw1_df$vec_char)
as.numeric(hw1_df$vec_factor)
```

As we can observe, the logical variable and the factor variable can be coverted and displayed in numerical format. However, the character variable cannot be coerced since it cannot be classified as the other two do. "0" and "1" can replace "false" and "true" messege in the logical variable while the content in factor variable can be placed into 10 different types and displayed in 10 numbers.

```{r convert character variable, results='hide', warning=FALSE}
as.character(hw1_df$vec_char)
as.numeric(hw1_df$vec_char)
as.factor(hw1_df$vec_char)
as.logical(hw1_df$vec_char)
```

The character variable can be converted into factor variable and be placed in a sequence as it has different levels. But it cannot be converted into numeric or logical variable since it does not contain numbers and it cannot be classified into two types either.

```{r convert factor variable, results='hide', warning=FALSE}
as.factor(hw1_df$vec_factor)
as.numeric(hw1_df$vec_factor)
as.character(hw1_df$vec_factor)
as.logical(hw1_df$vec_factor)
```

Similarly, the factor variable cannot be classified into two types, as a result, it cannot be converted into logical variable. It can be converted into numeric variable and shown in numbers according to the alphabetic order. It can also be converted into character variable since it is exactly a group of words.

